---
title: "Problem 5: Computational Finance - Modelling Stock prices"
author: "Deepmalya Dutta"
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
output:
  pdf_document: 
    df_print: paged
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Following piece of code download the prices of TCS since 2007

```{r, warning=FALSE}
library(quantmod)
getSymbols('TCS.NS')
tail(TCS.NS)
```
Plot the adjusted close prices of TCS
```{r}
plot(TCS.NS$TCS.NS.Adjusted)
```

**Download the data of market index Nifty50**. The Nifty 50 index indicates how the over all market has done over the similar period.
```{r, warning=FALSE}
getSymbols('^NSEI')
tail(NSEI)
```
Plot the adjusted close value of Nifty50
```{r}
plot(NSEI$NSEI.Adjusted)
```


### Log-Return 
We calculate the daily log-return, where log-return is defined as
$$
r_t=\log(P_t)-\log(P_{t-1})=\Delta \log(P_t),
$$
where $P_t$ is the closing price of the stock on $t^{th}$ day.

```{r}
TCS_rt = diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt = diff(log(NSEI$NSEI.Adjusted))
retrn = cbind.xts(TCS_rt,Nifty_rt) 
retrn = na.omit(data.frame(retrn))

plot(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted
     ,pch=20
     ,xlab='Market Return'
     ,ylab='TCS Return'
     ,xlim=c(-0.18,0.18)
     ,ylim=c(-0.18,0.18))
grid(col='grey',lty=1)
```

+ Consider the following model:

$$
r_{t}^{TCS}=\alpha + \beta r_{t}^{Nifty} + \varepsilon,
$$
where $\mathbb{E}(\varepsilon)=0$ and $\mathbb{V}ar(\varepsilon)=\sigma^2$.

1. Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$ using the method of moments type plug-in estimator discussed in the class.

```{r, warning = F}
m_x <- mean(retrn$NSEI.Adjusted)
m_y <- mean(retrn$TCS.NS.Adjusted)
s_x <- sd(retrn$NSEI.Adjusted)
s_y <- sd(retrn$TCS.NS.Adjusted)
r <- cor(retrn$TCS.NS.Adjusted, retrn$NSEI.Adjusted)

a0 <- m_y -r*s_y/s_x*m_x
b0 <- r*s_y/s_x

return <- retrn
return$pred_TCS <- a0 + b0*return$NSEI.Adjusted
e <- return$TCS.NS.Adjusted - return$pred_TCS
sigma0 <- sd(e)

print(c("The parameters of the model estimated using method of moments are:", paste("alpha = ", a0,", beta = ", b0, ", sigma =", sigma0)))
```

2. Estimate the parameters using the `lm` built-in function of `R`. Note that `lm` using the OLS method.
```{r, warning = F}
ols <- lm(TCS.NS.Adjusted~NSEI.Adjusted,data = retrn)
a1 <- ols$coefficients[1]
b1 <- ols$coefficients[2]
sigma1 <- sd(ols$residuals)

print(c("The parameters of the model estimated using `lm` built-in function of `R` are:", paste("alpha = ", a1,", beta = ", b1, ", sigma =", sigma1)))
```
3. Fill-up the following table

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |                   |
$\beta$    |                   |
$\sigma$   |                   |

#### Solution:

Parameters |  Method of Moments     |       OLS
-----------|------------------------|---------------------
$\alpha$   |  0.000462823035007132  | 0.000462823035007131
$\beta$    |  0.743684314730296     | 0.743684314730293
$\sigma$   |  0.0161846620846538    | 0.0161846620846538


4. If the current value of Nifty is 18000 and it goes up to 18200. The current value of TCS is Rs. 3200/-. How much you can expect TCS price to go up?

```{r, warning = F}
nifty_current <- 18000
nifty_future <- 18200
TCS_current <- 3200
nifty_return <- log(nifty_future) - log(nifty_current)
TCS_pred_return <- predict(ols, data.frame(NSEI.Adjusted = c(nifty_return)))
print(paste("The TCS price would go up to", round(exp(TCS_pred_return)*TCS_current)))
```