---
title: "Question2"
output:
  pdf_document: default
  html_document: default
date: "Varun Agrawal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(ggplot2)
library(cowplot)
library(dplyr)
library(reshape2)
library(latex2exp)
```

## Problem 2 : Simulation Study to Understand Sampling Distribution

**Part A**
Suppose $X_1,X_2,\cdots,X_n\stackrel{iid}{\sim} Gamma(\alpha,\sigma)$, with pdf as
$$
f(x | \alpha,\sigma)=\frac{1}{\sigma^{\alpha}\Gamma(\alpha)}e^{- x/\sigma}x^{\alpha-1},~~~~0<x<\infty,
$$
The mean and variance are $E(X)=\alpha\sigma$ and $Var(X)=\alpha\sigma^2$. Note that `shape = ` $\alpha$ and `scale = ` $\sigma$.

1. Write a `function` in `R` which will compute the MLE of $\theta=\log(\alpha)$ using `optim` function in `R`. You can name it `MyMLE`
```{r, warning=FALSE,message=FALSE}

MyMLE <- function(x,alfa,beta){
  est <- function(x1 = x,para){
    est = -1*sum(dgamma(x1,shape = para[1], scale = para[2],log = T))
    return(est)
    }
  opt <- optim(par = c(alfa,beta), fn = est,x1 = x)
  return(log(opt$par[1]))
}
```
```{r, warning=F,message=F}
Q2_1 <- function(n,N,alfa,beta){
  mle <- c(rep(0,N))
  for(i in 1:N){
    x <- rgamma(n,shape = alfa, scale = beta)
    mle[i] = MyMLE(x,alfa,beta)
  }
  q2_5<- as.character(round(quantile(mle, 0.025),3))
  q97_5<- as.character(round(quantile(mle,0.975),3))
  diff <- as.character(round((quantile(mle,0.975) - quantile(mle, 0.025)),3)) 
  
  hist(mle, col = 'skyblue3', breaks = 25)
  abline(v = log(alfa),lwd = 1.5, col = 'red')
  s1 <- as.character(round(log(alfa),2))
  lbl_str <- paste('True Value', s1)
  text(log(alfa),-1,lbl_str,col = 'red')
  print(paste('2.5th quantile = ',q2_5))
  print(paste('97.5th quantile = ',q97_5))
  print(paste('97.5th qunatile - 2.5th quantile = ',diff))
  return(round((quantile(mle,0.975) - quantile(mle, 0.025)),3))
}

```


2. Choose `n=20`, and `alpha=1.5` and `sigma=2.2`
     (i) Simulate $\{X_1,X_2,\cdots,X_n\}$ from `rgamma(n=20,shape=1.5,scale=2.2)`
     (ii) Apply the `MyMLE` to estimate $\theta$ and append the value in a vector
     (iii) Repeat the step (i) and (ii) 1000 times
     (iv) Draw histogram of the estimated MLEs of $\theta$.
     (v) Draw a vertical line using `abline` function at the true value of $\theta$.
     (vi) Use `quantile` function on estimated $\theta$'s to find the 2.5 and 97.5-percentile points. 
```{r}
q2 <- Q2_1(20,1000,1.5,2.2)
```
3.  Choose `n=40`, and `alpha=1.5` and repeat the (2).
```{r}
q3 <- Q2_1(40,1000,1.5,2.2)
```

4.  Choose `n=100`, and `alpha=1.5` and repeat the (2).
```{r}
q4 <- Q2_1(100,1000,1.5,2.2)
```

5. Check if the gap between 2.5 and 97.5-percentile points are shrinking as sample size `n` is increasing?
```{r}
print(q2>q3 && q3>q4)
```
Hence we can see that the gap between 2.5 and 97.5-percentile points are shrinking as sample size `n` is increasing. 


*Hint*: Perhaps you should think of writing a single `function` where you will provide the values of `n`, `sim_size`, `alpha` and `sigma`; and it will return the desired output. 
