---
title: "Question3"
output:
  html_document: default
  pdf_document: default
date: "Varun Agrawal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(ggplot2)
library(cowplot)
library(dplyr)
library(reshape2)
library(latex2exp)
```

## Problem 3: Analysis of `faithful` datasets.

Consider the `faithful` datasets:
```{r}
attach(faithful)
hist(faithful$waiting,xlab = 'waiting',probability = T,col='pink',main='')
```

Fit following three models using MLE method and calculate **Akaike information criterion** (aka., AIC) for each fitted model. Based on AIC decides which model is the best model? Based on the best model calculate the following probability
$$
\mathbb{P}(60<\texttt{waiting}<70)
$$

(i) **Model 1**:
$$
f(x)=p*Gamma(x|\alpha,\sigma_1)+(1-p)N(x|\mu,\sigma_2^2),~~0<p<1
$$

```{r, warning=FALSE,message=FALSE}
attach(faithful)
data=faithful
waiting=sort(waiting)
Myopt1 <- function(x,p,alfa,sigma1,mu,sigma2){
  est <- function(x1 = x,para){
    est = (-sum(log(para[1]*dgamma(x,para[2],para[3]) + (1-para[1])*dnorm(x,para[4],para[5]))))
    return(est)
    }
  opt <- optim(par = c(p,alfa,sigma1,mu,sigma2),control=list(maxit=10000), fn = est,x1 = x)
  aic_1 <- 10 + 2*est(x1 = x,para = opt$par)
  print(paste("AIC for model 1 = ",aic_1))
  return(c(opt$par,aic_1))
}
p = length(waiting[waiting<70])/length(waiting)
MyFit <- Myopt1(waiting,p,55,1,85,1)
AIC_1 <- Myopt1(waiting,p,55,1,85,1)[6]
#Estimators
phat <- MyFit[1]
alfahat <- MyFit[2]
sigma1hat <- MyFit[3]
muhat <- MyFit[4]
sigma2hat <- MyFit[5]
dplot=(phat*dgamma(waiting,alfahat,sigma1hat))+((1-phat)*dnorm(waiting,muhat,sigma2hat))

hist(waiting,probability = T,breaks = 25,col = 'skyblue3')
lines(waiting,dplot,lwd=2,col="red")

```


(ii) **Model 2**:
$$
f(x)=p*Gamma(x|\alpha_1,\sigma_1)+(1-p)Gamma(x|\alpha_2,\sigma_2),~~0<p<1
$$

```{r, warning=FALSE,message=FALSE}

attach(faithful)
data=faithful
waiting=sort(waiting)
Myopt2 <- function(x,p,alfa,sigma1,mu,sigma2){
  est <- function(x1 = x,para){
    est = (-sum(log(para[1]*dgamma(x,para[2],para[3]) + (1-para[1])*dgamma(x,para[4],para[5]))))
    return(est)
    }
  opt <- optim(par = c(p,alfa,sigma1,mu,sigma2),control=list(maxit=10000), fn = est,x1 = x)
  aic_2 <- 10 + 2*est(x1 = x,para = opt$par)
  print(paste("AIC for model 1 = ",aic_2))
  return(c(opt$par,aic_2))
}
p = length(waiting[waiting<70])/length(waiting)
MyFit2 <- Myopt2(waiting,p,55,1,85,1)
AIC_2 <- Myopt2(waiting,p,55,1,85,1)[6]
#Estimators
phat2 <- MyFit2[1]
alfahat2 <- MyFit2[2]
sigma1hat2 <- MyFit2[3]
muhat2 <- MyFit2[4]
sigma2hat2 <- MyFit2[5]
dplot2=(phat2*dgamma(waiting,alfahat2,sigma1hat2))+((1-phat2)*dgamma(waiting,muhat2,sigma2hat2))

hist(waiting,probability = T,breaks = 25,col = 'skyblue3')
lines(waiting,dplot2,lwd=2,col="red")
```
(iii) **Model 3**:
$$
f(x)=p*logNormal(x|\mu_1,\sigma_1^2)+(1-p)logNormal(x|\mu_1,\sigma_1^2),~~0<p<1
$$

```{r, warning=FALSE,message=FALSE}
attach(faithful)
data=faithful
waiting=sort(waiting)
Myopt3 <- function(x,p,alfa,sigma1,mu,sigma2){
  est <- function(x1 = x,para){
    est = (-sum(log(para[1]*dlnorm(x,para[2],para[3]) + (1-para[1])*dlnorm(x,para[4],para[5]))))
    return(est)
    }
  opt <- optim(par = c(p,alfa,sigma1,mu,sigma2),control=list(maxit=10000), fn = est,x1 = x)
  aic_3 <- 10 + 2*est(x1 = x,para = opt$par)
  print(paste("AIC for model 1 = ",aic_3))
  return(c(opt$par,aic_3))
}
p = length(waiting[waiting<70])/length(waiting)
MyFit3 <- Myopt3(waiting,p,3.9,0.015,4.4,0.012)
AIC_3 <- Myopt3(waiting,p,3.9,0.015,4.4,0.012)[6]
#Estimators
phat3 <- MyFit3[1]
alfahat3 <- MyFit3[2]
sigma1hat3 <- MyFit3[3]
muhat3 <- MyFit3[4]
sigma2hat3 <- MyFit3[5]
dplot3=(phat3*dlnorm(waiting,alfahat3,sigma1hat3))+((1-phat3)*dlnorm(waiting,muhat3,sigma2hat3))

hist(waiting,probability = T,breaks = 25,col = 'skyblue3')
lines(waiting,dplot3,lwd=2,col="red")
```

```{r,warning=FALSE,message=FALSE}
listAIC <- c('AIC1' = AIC_1,'AIC2' = AIC_2,'AIC3' = AIC_3)
minAIC <- which.min(listAIC)
minAIC
```

As we can see AIC for 3$^{rd}$ model is lowest So we will use it for getting our required probability, $$\mathbb{P}(60<\texttt{waiting}<70)$$


```{r,warning=FALSE,message=FALSE}
reqd_prob <- (phat3*plnorm(70, meanlog = alfahat3, sdlog = sigma1hat3) + (1-phat3)*plnorm(70, meanlog = muhat3, sdlog = sigma2hat3)) - (phat3*plnorm(60, meanlog = alfahat3, sdlog = sigma1hat3) + (1-phat3)*plnorm(60, meanlog = muhat3, sdlog = sigma2hat3))
print(paste('Require Prob is = ', reqd_prob))
```